---
title: "CORE Pipeline Quality Report"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    code-fold: true
    # embed-resources: false so all 109 renders share a single bootstrap +
    # icon-font asset dir under docs/quality-reports/_libs/. Phase 7
    # post-processes to dedupe and rewrite per-file references. With
    # embed-resources: true each report bloats to ~1.95 MB (mostly an inlined
    # Source Sans Pro TTF); with sharing, each is ~51 KB.
    embed-resources: false
params:
  report_data_path: "data/logs/quality_990ez_2017.rds"
---

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(data.table)
  library(knitr)
})

r <- readRDS(params$report_data_path)

fmt_n  <- function(n) if (is.na(n) || is.null(n)) "—" else format(n, big.mark = ",", scientific = FALSE)
fmt_pct <- function(p) if (is.na(p) || is.null(p)) "—" else sprintf("%.2f%%", p)
fmt_money <- function(v) if (is.na(v) || is.null(v)) "—" else sprintf("$%s", format(round(v), big.mark = ",", scientific = FALSE))

status_badge <- function(passed) if (isTRUE(passed)) "✅ PASSED" else "❌ FAILED"
```

---

**Form:** `r r$form`  **Tax year:** `r r$tax_year`  **Generated:** `r format(r$timestamp, "%Y-%m-%d %H:%M:%S")`

`r status_badge(r$hard_passed)`  — Strict mode: `r r$strict_mode`

## Executive Summary

| Metric | Value |
|---|---|
| Row count | `r fmt_n(r$row_count)` |
| Column count | `r r$column_count` |
| Vintage-aware completeness | `r fmt_pct(r$overall_completeness)` |
| Raw completeness (all cols / all rows) | `r fmt_pct(r$overall_completeness_raw)` |
| Unique EINs | `r fmt_n(r$summary_stats$unique_eins)` |
| Duplicate EINs (amendments) | `r fmt_n(r$summary_stats$duplicate_eins)` |
| Missing columns vs schema | `r length(r$missing_columns)` |
| Extra columns vs schema | `r length(r$extra_columns)` |
| YoY row-count delta vs prior run | `r if (r$summary_stats$year_over_year_delta$status == "ok") sprintf("%+0.2f%% (within ±20%%)", r$summary_stats$year_over_year_delta$delta_pct) else r$summary_stats$year_over_year_delta$status` |

## Identity-Field Checks

```{r identity-checks}
ein  <- r$critical_field_issues$ein
tp   <- r$critical_field_issues$tax_period
sub  <- r$critical_field_issues$subsection
typ  <- r$critical_field_issues$types

ident <- data.table(
  Check = c("EIN format (XX-XXXXXXX)", "EIN null count", "EIN duplicates (soft)",
            "tax_period format (YYYYMM)", "tax_period out-of-range",
            "subsection_cd whitelist", "Column types"),
  Result = c(
    if (is.null(ein))  "✅ all valid" else sprintf("❌ %d malformed", ein$malformed),
    if (is.null(ein) && r$summary_stats$duplicate_eins >= 0) "✅ no nulls" else sprintf("❌ %d nulls", ein$null_count),
    sprintf("ℹ %d (soft, not a failure)", r$summary_stats$duplicate_eins),
    if (is.null(tp))   "✅ all valid"  else sprintf("❌ %d malformed", tp$malformed_count),
    if (is.null(tp))   "✅ all in range" else sprintf("❌ %d out of range", tp$out_of_range_count),
    if (is.null(sub))  "✅ all in IRM whitelist" else sprintf("❌ %d unknown codes", sub$unknown_count),
    if (is.null(typ))  "✅ all types correct" else sprintf("❌ %s", paste(typ$issues, collapse = "; "))
  )
)
kable(ident)
```

## Subsection Code Distribution

(IRC 501(c) / 501(d) / 501(e) / 501(f) / 501(k) / 501(n) / 521 / 527 / 529 / 4947 / 1381 subsection codes per IRM 25.7.1. See `data/lookups/subsection_codes.csv`.)

```{r subsection-dist}
sd <- r$summary_stats$subsection_distribution
if (length(sd$top_values) == 0L) {
  cat("(no subsection data)")
} else {
  tab <- rbindlist(lapply(sd$top_values, as.data.table))
  setnames(tab, c("code", "count", "pct"))
  tab[, pct := paste0(pct, "%")]
  tab[, count := format(count, big.mark = ",")]
  kable(tab)
}
```

## Financial Summary

```{r financial}
fin <- r$summary_stats$financial
if (length(fin) == 0L) {
  cat("(no financial summary)")
} else {
  tab <- data.table(
    metric = c("total_revenue", "total_expenses", "total_assets_eoy"),
    sum = c(fmt_money(fin$revenue$total),  fmt_money(fin$expenses$total),  fmt_money(fin$assets$total)),
    median = c(fmt_money(fin$revenue$median), fmt_money(fin$expenses$median), fmt_money(fin$assets$median))
  )
  kable(tab)
}
```

## Tax Period Year Coverage

```{r tax-period-coverage}
d <- r$summary_stats$tax_period_year_distribution
if (length(d) == 0L) {
  cat("(no coverage data)")
} else {
  tab <- data.table(tax_year = names(d), n = as.integer(unlist(d)))
  setorder(tab, tax_year)
  tab[, pct := sprintf("%.2f%%", 100 * n / sum(n))]
  tab[, n := format(n, big.mark = ",")]
  kable(tab)
}
```

## Field Completeness by Category

Columns are grouped by Form section parsed from the crosswalk's `location` field.

```{r category-completeness}
cats <- r$category_reports
if (length(cats) == 0L) {
  cat("(no category breakdown)")
} else {
  tab <- rbindlist(lapply(cats, function(c) data.table(
    category         = c$category_name,
    n_cols           = c$column_count,
    cols_present     = c$columns_present,
    avg_completeness = sprintf("%.2f%%", c$avg_completeness)
  )))
  setorder(tab, -n_cols)
  kable(tab)
}
```

## Completeness by Vintage Cohort

When a single tax-year file blends rows from multiple `extract_year` × `source_form` cohorts, the per-cohort completeness reveals whether any vintage is dragging the overall metric down. A clean run produces 100% for every cohort.

```{r per-cohort}
pc <- r$completeness_by_cohort
if (length(pc) == 0L) {
  cat("(single cohort)")
} else {
  tab <- rbindlist(lapply(pc, function(c) data.table(
    extract_year     = c$extract_year,
    source_form      = c$source_form,
    rows             = format(c$n_rows, big.mark = ","),
    expected_cols    = c$n_expected_cols,
    completeness_pct = sprintf("%.2f%%", c$completeness_pct)
  )))
  kable(tab)
}
```

## Data Issues

```{r data-issues}
issues <- character(0)
if (length(r$missing_columns)) issues <- c(issues, sprintf("Missing %d columns: %s",
  length(r$missing_columns), paste(head(r$missing_columns, 10), collapse = ", ")))
if (length(r$extra_columns)) issues <- c(issues, sprintf("Extra %d columns: %s",
  length(r$extra_columns), paste(head(r$extra_columns, 10), collapse = ", ")))
if (length(r$critical_field_issues)) {
  for (k in names(r$critical_field_issues)) {
    issues <- c(issues, sprintf("%s: %s", k, paste(unlist(r$critical_field_issues[[k]]), collapse = " | ")))
  }
}

if (length(issues) == 0L) {
  cat("✅ No hard-check issues detected.\n")
} else {
  for (i in issues) cat("- ", i, "\n", sep = "")
}
```

---

*Generated from `r params$report_data_path`. See `R/quality/post_checks.R` for the validator implementation and `docs/05-quality-gates.qmd` for the gate definitions.*
