---
title: "BMF Pipeline Quality Report"
subtitle: "Post-Transformation Validation Results"
date: "`r Sys.time()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: 2
    theme: cosmo
    code-fold: true
    self-contained: true
    smooth-scroll: true
  pdf:
    toc: true
    toc-depth: 2
params:
  report_data_path: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)

library(ggplot2)
library(data.table)
library(scales)
library(knitr)

# Load report data from RDS file
if (!is.null(params$report_data_path) && file.exists(params$report_data_path)) {
  report <- readRDS(params$report_data_path)
} else {
  stop("Report data not found. Please provide a valid report_data_path parameter.")
}

# Helper function to format numbers
fmt_num <- function(x) format(x, big.mark = ",", scientific = FALSE)

# Helper function to format currency
fmt_currency <- function(x) paste0("$", format(x, big.mark = ",", scientific = FALSE))

# Helper function for status color
status_color <- function(pct) {
  if (is.na(pct)) return("#6c757d")
  if (pct >= 95) return("#28a745")
  if (pct >= 70) return("#ffc107")
  return("#dc3545")
}

# Helper function for status label
status_label <- function(pct) {
  if (is.na(pct)) return("N/A")
  if (pct >= 95) return("Good")
  if (pct >= 70) return("Fair")
  return("Low")
}
```

# Executive Summary

```{r summary-badge, results='asis'}
overall_color <- if (report$passed) "#28a745" else "#dc3545"
overall_text <- if (report$passed) "PASSED" else "FAILED"
overall_icon <- if (report$passed) "check-circle" else "x-circle"

cat(sprintf('
::: {.callout-%s appearance="simple"}
## Overall Status: %s

**Pipeline Timestamp:** %s
:::
', if(report$passed) "tip" else "warning", overall_text, format(report$timestamp, "%Y-%m-%d %H:%M:%S")))
```

```{r summary-metrics}
summary_df <- data.frame(
  Metric = c(
    "Total Records",
    "Output Columns",
    "Overall Completeness",
    "Row Preservation",
    "Unique EINs",
    "Duplicate EINs"
  ),
  Value = c(
    fmt_num(report$row_count),
    as.character(report$column_count),
    sprintf("%.1f%%", report$overall_completeness),
    if (report$row_preservation) "Passed" else "Failed",
    if (!is.null(report$summary_stats$unique_eins)) fmt_num(report$summary_stats$unique_eins) else "N/A",
    if (!is.null(report$summary_stats$duplicate_eins) && report$summary_stats$duplicate_eins > 0)
      fmt_num(report$summary_stats$duplicate_eins) else "0"
  )
)

kable(summary_df, col.names = c("Metric", "Value"), align = c("l", "r"))
```

---

# Field Completeness by Category

This section shows the completeness of each output field, grouped by category. Each field shows its source column(s) from the raw BMF file.

```{r completeness-setup}
# Build completeness dataframe from column_completeness
if (!is.null(report$column_completeness) && length(report$column_completeness) > 0) {
  completeness_list <- report$column_completeness

  completeness_df <- data.frame(
    Field = names(completeness_list),
    Source = sapply(completeness_list, function(x) {
      src <- x$source_columns
      if (length(src) > 1) paste(src, collapse = ", ") else as.character(src)
    }),
    Completeness = sapply(completeness_list, function(x) x$completeness_pct),
    NullCount = sapply(completeness_list, function(x) x$null_count),
    EmptyCount = sapply(completeness_list, function(x) x$empty_count),
    stringsAsFactors = FALSE,
    row.names = NULL
  )
}
```

```{r completeness-chart, fig.cap="Fields with Incomplete Data", fig.height=8}
if (exists("completeness_df") && nrow(completeness_df) > 0) {
  incomplete_df <- completeness_df[completeness_df$Completeness < 100, ]

  if (nrow(incomplete_df) > 0) {
    incomplete_df <- incomplete_df[order(incomplete_df$Completeness), ]
    incomplete_df$Field <- factor(incomplete_df$Field, levels = incomplete_df$Field)

    ggplot(incomplete_df, aes(x = Field, y = Completeness)) +
      geom_col(aes(fill = Completeness), width = 0.7) +
      geom_hline(yintercept = 95, linetype = "dashed", color = "#28a745", linewidth = 0.8) +
      geom_hline(yintercept = 70, linetype = "dashed", color = "#ffc107", linewidth = 0.8) +
      geom_text(aes(label = sprintf("%.1f%%", Completeness)), hjust = -0.1, size = 3) +
      scale_fill_gradient2(
        low = "#dc3545",
        mid = "#ffc107",
        high = "#28a745",
        midpoint = 82.5,
        limits = c(0, 100)
      ) +
      coord_flip() +
      labs(
        title = "Fields with Less Than 100% Completeness",
        subtitle = "Green line = 95% (Good), Yellow line = 70% (Fair threshold)",
        x = NULL,
        y = "Completeness (%)"
      ) +
      theme_minimal() +
      theme(
        legend.position = "none",
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 9)
      ) +
      scale_y_continuous(limits = c(0, 105), breaks = seq(0, 100, 20))
  } else {
    cat("\n\n**All fields have 100% completeness.**\n\n")
  }
}
```

```{r category-tables, results='asis'}
# Define category order and display names
category_info <- list(
  identity = "Identity",
  org_name = "Organization Name",
  dba_name = "DBA Name",
  ico = "In Care Of",
  group_exemption = "Group Exemption",
  address_raw = "Address (Raw)",
  address_clean = "Address (Cleaned)",
  address_flags = "Address Quality Flags",
  classification = "Classification",
  codes = "Organization Codes",
  dates = "Dates",
  financial_codes = "Financial Codes",
  financial_amounts = "Financial Amounts",
  activity = "Activity",
  filing = "Filing Requirements",
  ntee = "NTEE Codes",
  nteev2 = "NTEE V2 Codes"
)

if (!is.null(report$category_reports)) {
  for (cat_key in names(category_info)) {
    cat_report <- report$category_reports[[cat_key]]
    if (is.null(cat_report)) next

    cat_name <- category_info[[cat_key]]
    avg_comp <- cat_report$avg_completeness

    # Section header with completeness badge
    status_class <- if (avg_comp >= 95) "tip" else if (avg_comp >= 70) "warning" else "important"

    cat(sprintf('\n\n## %s {.unnumbered}\n\n', cat_name))
    cat(sprintf('**Average Completeness:** %.1f%% | **Columns:** %d\n\n',
                avg_comp, cat_report$column_count))

    # Build table for this category
    cat_cols <- names(cat_report$columns)
    if (length(cat_cols) > 0) {
      cat_data <- data.frame(
        Field = character(),
        Source = character(),
        Completeness = character(),
        Status = character(),
        `Null Count` = character(),
        stringsAsFactors = FALSE,
        check.names = FALSE
      )

      for (col_name in cat_cols) {
        col_info <- cat_report$columns[[col_name]]
        if (!col_info$present) next

        comp_pct <- if (!is.null(col_info$completeness)) col_info$completeness$completeness_pct else NA
        null_ct <- if (!is.null(col_info$completeness)) col_info$completeness$null_count else NA

        # Get source from source_column_map directly
        src <- "N/A"
        if (!is.null(report$source_column_map[[col_name]])) {
          src_cols <- report$source_column_map[[col_name]]
          src <- if (length(src_cols) > 1) paste(src_cols, collapse = ", ") else as.character(src_cols)
        }

        cat_data <- rbind(cat_data, data.frame(
          Field = col_name,
          Source = src,
          Completeness = if (!is.na(comp_pct)) sprintf("%.1f%%", comp_pct) else "N/A",
          Status = status_label(comp_pct),
          `Null Count` = if (!is.na(null_ct)) fmt_num(null_ct) else "N/A",
          stringsAsFactors = FALSE,
          check.names = FALSE
        ))
      }

      if (nrow(cat_data) > 0) {
        print(kable(cat_data, align = c("l", "l", "r", "c", "r"), row.names = FALSE))

        # Print column descriptions as bullet points
        cat("\n\n**Column Descriptions:**\n\n")
        for (col_name in cat_cols) {
          desc <- report$column_descriptions[[col_name]]
          if (!is.null(desc)) {
            cat(sprintf("- **%s**: %s\n", col_name, desc))
          }
        }
      }
    }
  }
}
```

---

# Organization Distribution

## Exempt Organization Type

Distribution of organizations by exempt organization type (based on IRS subsection code).

```{r org-type-chart, fig.cap="Organizations by Exempt Type", fig.height=6}
if (!is.null(report$summary_stats$org_type_distribution)) {
  orgtype_dt <- as.data.table(report$summary_stats$org_type_distribution)

  if (nrow(orgtype_dt) > 0 && "exempt_organization_type" %in% names(orgtype_dt)) {
    setnames(orgtype_dt, old = c("exempt_organization_type", "N"),
             new = c("Type", "Count"), skip_absent = TRUE)

    # Take top 15
    orgtype_dt <- orgtype_dt[order(-Count)][1:min(15, .N)]
    orgtype_dt <- orgtype_dt[!is.na(Type) & Type != ""]

    # Truncate long names
    orgtype_dt[, DisplayName := ifelse(
      nchar(Type) > 50,
      paste0(substr(Type, 1, 47), "..."),
      Type
    )]

    orgtype_dt[, DisplayName := factor(DisplayName, levels = rev(DisplayName))]

    ggplot(orgtype_dt, aes(x = DisplayName, y = Count)) +
      geom_col(fill = "#4e79a7", width = 0.7) +
      geom_text(aes(label = fmt_num(Count)), hjust = -0.1, size = 3.5) +
      coord_flip() +
      scale_y_continuous(
        labels = label_comma(),
        expand = expansion(mult = c(0, 0.15))
      ) +
      labs(
        title = "Organizations by Exempt Type",
        subtitle = sprintf("Total: %s organizations | Top 15 shown", fmt_num(sum(orgtype_dt$Count))),
        x = NULL,
        y = "Number of Organizations"
      ) +
      theme_minimal() +
      theme(
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 9)
      )
  }
}
```

```{r org-type-table}
if (!is.null(report$summary_stats$org_type_distribution)) {
  orgtype_dt <- as.data.table(report$summary_stats$org_type_distribution)

  if (nrow(orgtype_dt) > 0 && "exempt_organization_type" %in% names(orgtype_dt)) {
    total <- sum(orgtype_dt$N)
    orgtype_dt[, Percentage := sprintf("%.2f%%", 100 * N / total)]
    orgtype_dt[, N := fmt_num(N)]
    setnames(orgtype_dt, c("Exempt Organization Type", "Count", "Percentage"))

    kable(orgtype_dt, align = c("l", "r", "r"))
  }
}
```

## NTEE Major Group Distribution

Distribution of organizations by NTEE major group code.

```{r ntee-chart, fig.cap="Organizations by NTEE Major Group", fig.height=8}
if (!is.null(report$summary_stats$ntee_major_group_distribution)) {
  ntee_dt <- as.data.table(report$summary_stats$ntee_major_group_distribution)

  if (nrow(ntee_dt) > 0) {
    setnames(ntee_dt, old = c("ntee_code_major_group", "N"),
             new = c("MajorGroup", "Count"), skip_absent = TRUE)

    ntee_dt <- ntee_dt[order(-Count)]
    ntee_dt[, MajorGroup := factor(MajorGroup, levels = rev(MajorGroup))]

    ggplot(ntee_dt, aes(x = MajorGroup, y = Count)) +
      geom_col(fill = "#59a14f", width = 0.7) +
      geom_text(aes(label = fmt_num(Count)), hjust = -0.1, size = 3.5) +
      coord_flip() +
      scale_y_continuous(
        labels = label_comma(),
        expand = expansion(mult = c(0, 0.15))
      ) +
      labs(
        title = "Organizations by NTEE Major Group",
        subtitle = sprintf("Total: %s organizations with NTEE codes", fmt_num(sum(ntee_dt$Count))),
        x = "NTEE Major Group",
        y = "Number of Organizations"
      ) +
      theme_minimal() +
      theme(
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 10, face = "bold")
      )
  }
}
```

```{r ntee-table}
if (!is.null(report$summary_stats$ntee_major_group_distribution)) {
  ntee_dt <- as.data.table(report$summary_stats$ntee_major_group_distribution)

  if (nrow(ntee_dt) > 0) {
    total <- sum(ntee_dt$N)
    ntee_dt[, Percentage := sprintf("%.2f%%", 100 * N / total)]
    ntee_dt[, N := fmt_num(N)]
    setnames(ntee_dt, c("NTEE Major Group", "Count", "Percentage"))

    kable(ntee_dt, align = c("l", "r", "r"))
  }
}
```

---

# Financial Summary

```{r financial-summary}
if (!is.null(report$summary_stats$financial)) {
  fin <- report$summary_stats$financial

  financial_df <- data.frame(
    Metric = character(),
    Value = character(),
    stringsAsFactors = FALSE
  )

  if (!is.null(fin$total_assets)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Total Assets (all organizations)",
      Value = fmt_currency(fin$total_assets)
    ))
  }

  if (!is.null(fin$median_assets)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Median Assets",
      Value = fmt_currency(fin$median_assets)
    ))
  }

  if (!is.null(fin$orgs_with_assets)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Organizations with Asset Data",
      Value = fmt_num(fin$orgs_with_assets)
    ))
  }

  if (!is.null(fin$orgs_zero_assets)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Organizations with Zero Assets",
      Value = fmt_num(fin$orgs_zero_assets)
    ))
  }

  if (!is.null(fin$total_income)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Total Income",
      Value = fmt_currency(fin$total_income)
    ))
  }

  if (!is.null(fin$median_income)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Median Income",
      Value = fmt_currency(fin$median_income)
    ))
  }

  if (!is.null(fin$total_revenue)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Total Revenue",
      Value = fmt_currency(fin$total_revenue)
    ))
  }

  if (!is.null(fin$median_revenue)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Median Revenue",
      Value = fmt_currency(fin$median_revenue)
    ))
  }

  if (nrow(financial_df) > 0) {
    kable(financial_df, col.names = c("Metric", "Value"), align = c("l", "r"))
  } else {
    cat("No financial summary data available.")
  }
} else {
  cat("No financial summary data available.")
}
```

---

# Address Quality

```{r address-quality}
if (!is.null(report$summary_stats$address_quality)) {
  aq <- report$summary_stats$address_quality

  address_df <- data.frame(
    Metric = c(
      "Missing Address",
      "P.O. Box Addresses",
      "Rural Route Addresses",
      "Invalid State Code"
    ),
    Count = c(
      fmt_num(aq$missing_count),
      fmt_num(aq$po_box_count),
      fmt_num(aq$rural_route_count),
      fmt_num(aq$invalid_state_count)
    ),
    Percentage = c(
      sprintf("%.2f%%", aq$missing_pct),
      sprintf("%.2f%%", aq$po_box_pct),
      sprintf("%.2f%%", 100 * aq$rural_route_count / report$row_count),
      sprintf("%.2f%%", 100 * aq$invalid_state_count / report$row_count)
    ),
    stringsAsFactors = FALSE
  )

  kable(address_df, col.names = c("Address Issue", "Count", "% of Total"), align = c("l", "r", "r"))
} else {
  cat("No address quality data available.")
}
```

---

# Date Coverage

## Ruling Date Range

```{r ruling-date}
if (!is.null(report$summary_stats$ruling_date_range)) {
  rdr <- report$summary_stats$ruling_date_range

  ruling_df <- data.frame(
    Metric = c(
      "Earliest Ruling Date",
      "Latest Ruling Date",
      "Organizations with Ruling Date"
    ),
    Value = c(
      as.character(rdr$earliest),
      as.character(rdr$latest),
      fmt_num(rdr$orgs_with_ruling_date)
    ),
    stringsAsFactors = FALSE
  )

  kable(ruling_df, col.names = c("Metric", "Value"), align = c("l", "r"))
} else {
  cat("No ruling date range data available.")
}
```

## Tax Period Year Distribution

```{r tax-period-chart, fig.cap="Organizations by Tax Period Year"}
if (!is.null(report$summary_stats$tax_period_year_distribution)) {
  tax_dt <- as.data.table(report$summary_stats$tax_period_year_distribution)

  if (nrow(tax_dt) > 0) {
    setnames(tax_dt, old = c("tax_year", "N"), new = c("Year", "Count"), skip_absent = TRUE)
    tax_dt <- tax_dt[order(Year)]

    ggplot(tax_dt, aes(x = Year, y = Count)) +
      geom_col(fill = "#e15759", width = 0.7) +
      geom_text(aes(label = fmt_num(Count)), vjust = -0.5, size = 3.5) +
      scale_y_continuous(
        labels = label_comma(),
        expand = expansion(mult = c(0, 0.1))
      ) +
      labs(
        title = "Organizations by Tax Period Year",
        x = "Tax Period Year",
        y = "Number of Organizations"
      ) +
      theme_minimal() +
      theme(
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)
      )
  }
}
```

```{r tax-period-table}
if (!is.null(report$summary_stats$tax_period_year_distribution)) {
  tax_dt <- as.data.table(report$summary_stats$tax_period_year_distribution)

  if (nrow(tax_dt) > 0) {
    total <- sum(tax_dt$N)
    tax_dt[, Percentage := sprintf("%.2f%%", 100 * N / total)]
    tax_dt[, N := fmt_num(N)]
    setnames(tax_dt, c("Tax Year", "Count", "Percentage"))

    kable(tax_dt, align = c("l", "r", "r"))
  }
}
```

---

# Data Issues

```{r missing-cols, results='asis'}
if (length(report$missing_columns) > 0) {
  cat("\n\n## Missing Expected Columns\n\n")
  cat("::: {.callout-warning}\n")
  cat("The following columns were expected but not found in the output:\n\n")
  for (col in report$missing_columns) {
    cat(sprintf("- `%s`\n", col))
  }
  cat(":::\n")
}
```

```{r critical-issues, results='asis'}
if (length(report$critical_field_issues) > 0) {
  cat("\n\n## Critical Field Issues\n\n")
  cat("::: {.callout-important}\n")
  cat("The following critical fields have NULL or empty values:\n\n")
  for (field in names(report$critical_field_issues)) {
    cat(sprintf("- **%s**: %s NULL/empty values\n",
                field,
                fmt_num(report$critical_field_issues[[field]])))
  }
  cat(":::\n")
}
```

```{r extra-cols, results='asis'}
if (length(report$extra_columns) > 0) {
  cat("\n\n## Extra Columns (Not in Expected Schema)\n\n")
  cat("::: {.callout-note}\n")
  cat("The following columns were found but not in the expected output schema:\n\n")
  for (col in report$extra_columns) {
    cat(sprintf("- `%s`\n", col))
  }
  cat(":::\n")
}
```

```{r no-issues, results='asis'}
if (length(report$missing_columns) == 0 &&
    length(report$critical_field_issues) == 0 &&
    length(report$extra_columns) == 0) {
  cat("\n\n::: {.callout-tip}\n")
  cat("## No Data Issues Detected\n\n")
  cat("All expected columns are present and no critical field issues were found.\n")
  cat(":::\n")
}
```

---

# Report Metadata

```{r metadata}
meta_df <- data.frame(
  Property = c(
    "Report Generated",
    "Pipeline Timestamp",
    "Row Preservation Check",
    "Overall Completeness",
    "Overall Result"
  ),
  Value = c(
    format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
    format(report$timestamp, "%Y-%m-%d %H:%M:%S"),
    if (report$row_preservation) "Passed" else "Failed",
    sprintf("%.1f%%", report$overall_completeness),
    if (report$passed) "PASSED" else "FAILED"
  )
)

kable(meta_df, col.names = c("Property", "Value"), align = c("l", "l"))
```

---

*Generated by BMF Pipeline Quality System*
